#440 Reproducibility and Statistics Assignment Homework

<!-- This is Amanda Walker's Assignment. -->
##Amanda Walker, PSY441 EOU Spring 2015

<!-- Loaded politics.csv data file. -->

```{r}
politics<-read.csv("politics.csv")
```

###Original Data Structure

```{r, echo=FALSE}
str(politics)
```

<!-- Restructured the data for factors and proper variables -->

Restructure

```{r}
politics$subject<-factor(politics$subject)
politics$testtime<-factor(politics$testtime, levels=c("pre", "post"))
str(politics)
```

<!-- The minimum, mean, and variance for posttest
optimism scores. -->

###Summary of posttest optimism scores

```{r}
summary(politics$optimismscore[politics$testtime=="post"])
```

<!-- A histogram of posttest optimism scores with the default ugly settings. -->

###Histogram of posttest optimism scores

```{r}
hist(politics$optimismscore[politics$testtime=="post"])
```

<!-- A test to determine whether political affiliation
is independent of one's gender in tabular form.
-->

###Test of political affiliation and sex correlation

Frequencies

```{r}
pre=(politics[TRUE,])
post=(politics[!TRUE,])
tab<-table(pre$party,pre$sex)
tab
```

ChiSquare test

```{r}
chisq.test(pre$party,pre$sex)
```

Political affiliation is independent of sex, Chi-Square(2)=1.4535, p=0.4835

<!-- A test to determine whether or not males and females
have different incomes and a graph to display data that includes the relevant means and standard errors.-->

###Comparison of sex and income

```{r}
t.test(politics$income[politics$sex=="male"&politics$testtime=="post"],politics$income[politics$sex=="female"&politics$testtime=="post"])
```

Females demonstrated a weak association with higher incomes t(61.623)= -1.5714, p= 0.1212.

```{r}
library("dplyr")
temp<-politics[politics$testtime=="pre",]%>%group_by(sex)%>%summarize(means=mean(income),sems=sd(income)/sqrt(length(income)))
library("ggplot2")
f<-ggplot(temp,aes(x=factor(sex),y=means))+
    geom_bar(stat="identity", color="black",fill=c("deeppink","deepskyblue2"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems), width=.1)
f
```

<!--graph-->

```{r}
f<-f+ggtitle("Income by Sex")+
    labs(x="Sex", y="Income (thousand)")+
    scale_x_discrete(breaks=c("female","male"),labels=c("Female","Male"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(), axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())
f
```

<!-- ANOVA to see if male and females with different political affiliations
have different posttest optimism scores and a graph to display data that includes the relevant means and standard errors.-->

###Comparison of posttest optimism scores among males and females of different political affiliations

```{r}
 summary(aov(optimismscore~party*sex,data=politics[politics$testtime=="post",]))
```

Although optimism scores were related to political affiliation, F(2, 63) = 27.062, p < .001, they were not affected by sex, F(1, 63) = 0.040, p = 0.843, and there was no interaction between party and sex, F(2, 63) = 1.213, p = .304.

```{r}
temp<-politics[politics$testtime=="post",]%>%group_by(party,sex)%>%
    summarize(means=mean(optimismscore),sems=sd(optimismscore)/sqrt(length(optimismscore)))
library("gplots")
col1=col2hex("deeppink")
col2=col2hex("deepskyblue2")
f<-ggplot(temp, aes(x=party, y=means, fill=sex))+
    geom_bar(stat="identity",position=position_dodge())+
    scale_fill_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    theme(legend.key=element_rect(color="black"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2,position=position_dodge(.9))+
    ggtitle("Optimism Scores by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Optimism Score")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.76,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```

<!-- A line graph for the same data with
error bars.-->
A line graph of this data

```{r}
f<-ggplot(temp, aes(x=party, y=means, group=sex, color=sex))+
    geom_line(size=1)+
    geom_point(size=2)+
    scale_color_manual(values=c(col1,col2),name="Sex",breaks=c("female","male"),labels=c("Female", "Male"))+
    geom_errorbar(aes(ymax=means+sems, ymin=means-sems),width=.2)+
    ggtitle("Optimism Scores by Sex and Political Affiliation")+
    labs(x="Political Party Affiliation",y="Optimism Score")+
    scale_x_discrete(breaks=c("democrat","independent","republican"),labels=c("Democrat","Independent","Republican"))+
    theme(plot.title=element_text(size=15,face="bold",vjust=.5))+
    theme(axis.title.x=element_text(size=12,face="bold",vjust=-.25))+
    theme(axis.title.y=element_text(size=12,face="bold",vjust=1))+
    theme(axis.text.x=element_text(size=10,face="bold",color="black"))+
    theme(axis.text.y=element_text(size=10,face="bold",color="black"))+
    coord_cartesian(ylim=c(min(temp$means)-2*max(temp$sems),max(temp$means)+2*max(temp$sems)))+
    theme(panel.border=element_blank(),axis.line=element_line())+
    theme(panel.grid.major.x=element_blank())+
    theme(panel.grid.major.y=element_line(color="darkgrey"))+
    theme(panel.grid.minor.y=element_blank())+
    theme(legend.position=c(.76,.76))+
    theme(legend.background=element_blank())+
    theme(legend.background=element_rect(color="black"))+
    theme(legend.title=element_blank())+
    theme(legend.title=element_text(size=12))+
    theme(legend.title.align=.5)+
    theme(legend.text=element_text(size=10,face="bold"))
f
```


<!-- A test to see if optimism scores pre- and post
watching videos depends on sex.-->

###Comparison of optimism scores before and after watching videos among sexes
```{r}
summary(aov(optimismscore~testtime*sex+Error(subject/testtime),data=politics))
```

Although optimism scores were associated with watching the videos, F(1, 64) = 41.299, p< 0.001, there was no association between optimism score and sex F(1, 64)= 0.119 p=0.713, and there was no interation between taking the test before or after watching videos and their sex, F(1, 64) = 0.079, p = 0.825.


<!-- A Multiple Regression Analysis to see
whether we can predict someone's posttest optimism
scores from their pretest optimism scores and party
affiliation.-->

###Multiple Regression Analysis of posttest optimism score based on pretest optimism score and party

```{r}
summary(lm(optimismscore[politics$testtime=="post"]~optimismscore[politics$testtime=="pre"]+party[politics$testtime=="pre"], data=politics))
```

